Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-11 paper/paper.tex
edit first sentence
edit subtitles in RW section

Music information retrieval (MIR) operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).

@@

\subsection*{Timbre modeling as music information retrieval (MIR)}
On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags,'' encoded as binary outputs within some predefined label space.
Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.

@@

\subsection*{Timbre modeling as music cognition research}
On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
Rather, each excerpt must be played back to multiple human listeners.
Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, restricting the size of the musical corpus under study.
Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in deep neural networks.
Nevertheless, while training artificial neurons is prone to statistical overfitting, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}, thus yielding a useful form of inductive bias for downstream machine listening applications.

@@

The standard Euclidean distance on the raw scattering coefficients provides a universal representation of timbre dissimilarity.
We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.

@@

In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
Assuming invariance to timbre with respect to pitch and nuance is in line with the ANSI definition of timbre \ml{vl, citation. reformuler tout le paragraphe}, it is thus taken in this work as a reasonable assumption.
Quantification of this invariance is left for further study.

>>>>

Music information retrieval (MIR) operates at two levels: symbolic and auditory \cite{downie2003mir}.
By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).

@@

\subsection*{Timbre modeling in music information retrieval (MIR)}
On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags,'' encoded as binary outputs within some predefined label space.
Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.

@@

\subsection*{Timbre modeling in music cognition research}
On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
Rather, each excerpt must be played back to multiple human listeners.
Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, restricting the size of the musical corpus under study.
Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in deep neural networks.
Nevertheless, while training artificial neurons is prone to statistical overfitting, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}, thus yielding a useful form of inductive bias for downstream machine listening applications.

@@

The standard Euclidean distance on the raw scattering coefficients provides an instrument-agnostic representation of timbre dissimilarity.
We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.

@@

%In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
%Assuming invariance to timbre with respect to pitch and nuance is in line with the ANSI definition of timbre \ml{vl, citation. reformuler tout le paragraphe}, it is thus taken in this work as a reasonable assumption.
%Quantification of this invariance is left for further study.
